---
title: "summary_nadaraya"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{summary_nadaraya}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadarayaWatson)
library(np)
library(microbenchmark)
```

## Introduction

This package provides an implementation of the Nadaraya-Watson kernel regression estimator using Rcpp. The Nadaraya-Watson estimator is a nonparametric regression estimator that estimates the conditional expectation of a response variable given a predictor variable. The estimator is based on a weighted average of the response variable values, where the weights are determined by a kernel function. The package also provides functions for estimating confidence intervals for the regression function using a bootstrap resampling approach.

## mathematical expression and how to implement the model

For a given point \( x \), the Nadaraya-Watson estimator of \( Y \) is defined as:

\[
\hat{m}(x) = \frac{\sum_{i=1}^n K\left(\frac{x - X_i}{h}\right) Y_i}{\sum_{i=1}^n K\left(\frac{x - X_i}{h}\right)}
\]

where:

- \( K \) is the **kernel function**, a smooth, symmetric function (such as a Gaussian, Epanechnikov, or uniform kernel) that assigns weights to observations based on their distance from \( x \).
- \( h \) is the **bandwidth** parameter, controlling the width of the kernel. A small \( h \) focuses more on points closer to \( x \), leading to a more localized (and possibly noisier) estimate, while a larger \( h \) smooths over a broader range of points.

I use Cpp in implementation of the model and the Cpp considers the R interface and export the results to R

## Comparision between R implementation and Cpp implementation

In R implementation I do not use fully optimized method because the function requires nested for loop and the function also requires some methods of manipulations within the loop so directly using vecotorization is not trivial and it is very slow. Most of the implementation is in matlab and the matlab is fast in this. Here, we can see the comparison between the R implementation and the Cpp implementation.

```{r, fig.width=8, fig.height=6, dpi=150}
# Generate example data
set.seed(42)
x <- sort(seq(0, 10, length.out = 100))
y <- sin(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 10, length.out = 100)
bandwidth <- 0.5
n_boot <- 500

# Predict with bootstrapping for confidence intervals
results <- nadaraya_watson_r(x, y, x_eval, bandwidth, n_boot, conf_level = 0.95)

# Extract the results
y_pred <- results$y_pred
lower <- results$lower
upper <- results$upper

# Plot the results with confidence intervals
plot(x, y, main = "Nonparametric Regression with Confidence Intervals", xlab = "x", ylab = "y")
lines(x_eval, y_pred, col = "blue", lwd = 2, lty = 1)
lines(x_eval, lower, col = "blue", lty = 2)
lines(x_eval, upper, col = "blue", lty = 2)
legend("topright", legend = c("Estimate", "Confidence Interval"), col = c("blue", "blue"), lty = c(1, 2), lwd = 2)
```

We can also compare the running time of the two packages that implementation of the same result. The result indicates that the Cpp implementation is much faster than the R implementation.

```{r}
# Define test inputs
set.seed(42)
x <- sort(runif(100, 0, 10))
y <- sin(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 10, length.out = 100)
bandwidth <- 0.5
n_boot <- 50
conf_level <- 0.95

# Benchmark the R and C++ implementations
benchmark_results <- microbenchmark(
  R_version = nadaraya_watson_r(x, y, x_eval, bandwidth, n_boot, conf_level),
  Cpp_version = nadaraya_watson(x, y, x_eval, bandwidth, n_boot, conf_level),
  times = 5  # Run each function 10 times for a robust comparison
)

# Print benchmark results
print(benchmark_results)

```

## Results

Here is an example that you can consider to compare the model and the expected outcome of the np package. We can see that the results are pretty much related to each other and the differences are not that much. Here is the cpp implementation

```{r, fig.width=8, fig.height=6, dpi=150}
# Generate example data
set.seed(42)
x <- sort(seq(0, 10, length.out = 100))
y <- sin(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 10, length.out = 100)
bandwidth <- 0.5
n_boot <- 500

# Predict with bootstrapping for confidence intervals
results <- nadaraya_watson(x, y, x_eval, bandwidth, n_boot, conf_level = 0.95)

# Extract the results
y_pred <- results$y_pred
lower <- results$lower
upper <- results$upper

# Plot the results with confidence intervals
plot(x, y, main = "Nonparametric Regression with Confidence Intervals", xlab = "x", ylab = "y")
lines(x_eval, y_pred, col = "blue", lwd = 2, lty = 1)
lines(x_eval, lower, col = "blue", lty = 2)
lines(x_eval, upper, col = "blue", lty = 2)
legend("topright", legend = c("Estimate", "Confidence Interval"), col = c("blue", "blue"), lty = c(1, 2), lwd = 2)
```

here is the R implementation

```{r, fig.width=8, fig.height=6, dpi=150}
# Generate example data
set.seed(42)
x <- sort(seq(0, 10, length.out = 100))
y <- cos(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 10, length.out = 100)
bandwidth <- 0.5
n_boot <- 500

# Predict with bootstrapping for confidence intervals
results <- nadaraya_watson_r(x, y, x_eval, bandwidth, n_boot, conf_level = 0.95)

# Extract the results
y_pred <- results$y_pred
lower <- results$lower
upper <- results$upper

# Plot the results with confidence intervals
plot(x, y, main = "Nonparametric Regression with Confidence Intervals", xlab = "x", ylab = "y")
lines(x_eval, y_pred, col = "blue", lwd = 2, lty = 1)
lines(x_eval, lower, col = "blue", lty = 2)
lines(x_eval, upper, col = "blue", lty = 2)
legend("topright", legend = c("Estimate", "Confidence Interval"), col = c("blue", "blue"), lty = c(1, 2), lwd = 2)
```


## Some Differences between this package and the np package

This package uses some kind of different optimization comparing with the np package since the np package uses other tricks in modeling and this package only follows from the textbook. Moreover, implementing a large pacakge is beyond the scope of this homework, so I just leave this to future works.

We can choose to generate the code for comparision:

```{r}
# Generate example data
set.seed(42)
x <- sort(runif(100, 0, 10))
y <- sin(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 10, length.out = 100)
bandwidth <- 0.5
```

Now we can consider the graph to compare the results of the Rcpp implementation with the np package:

```{r, fig.width=8, fig.height=6, dpi=150}
# Predict with Rcpp implementation 
results <- nadaraya_watson(x, y, x_eval, bandwidth, n_boot = 500, conf_level = 0.95) 
y_pred_cpp <- results$y_pred
lower_cpp <- results$lower
upper_cpp <- results$upper

# Predict with the np package (R implementation)
np_model <- npreg(bws = bandwidth, xdat = x, ydat = y, exdat = x_eval, 
                  ckertype = "gaussian")
y_pred_np <- predict(np_model)

# Plot the results with confidence intervals
plot(x, y, main = "Nonparametric Regression Comparison", xlab = "x", ylab = "y")
lines(x_eval, y_pred_cpp, col = "blue", lwd = 2, lty = 1)
lines(x_eval, lower_cpp, col = "blue", lty = 2)
lines(x_eval, upper_cpp, col = "blue", lty = 2)
lines(x_eval, y_pred_np, col = "red", lwd = 2, lty = 1)
legend("topright", legend = c("Rcpp Estimate", "Rcpp Confidence Interval", "np Package Estimate"), 
       col = c("blue", "blue", "red"), lty = c(1, 2, 1), lwd = c(2, 1, 2))
```

We can also plot the differences explcitly

```{r, fig.width=8, fig.height=6, dpi=150}
# Display a summary of the differences
diff <- y_pred_cpp - y_pred_np
summary(diff)

# Optionally, plot the difference between the two estimates
plot(x_eval, diff, type = "b", col = "purple", main = "Difference Between Rcpp and np Estimates",
     xlab = "x_eval", ylab = "Difference (Rcpp - np)", pch = 19)
abline(h = 0, col = "gray", lty = 2)
```

Now we can also see the diffrences with the R implementation and the np package

```{r}
# Generate example data
set.seed(42)
x <- sort(runif(100, 0, 3))
y <- cos(x) + sin(x) + rnorm(100, sd = 0.2)
x_eval <- seq(0, 3, length.out = 100)
bandwidth <- 0.5
n_boot <- 500
```

Now we can consider the graph to compare the results of the Rcpp implementation with the np package:

```{r, fig.width=8, fig.height=6, dpi=150}
# Predict with Rcpp implementation 
results <- nadaraya_watson_r(x, y, x_eval, bandwidth, n_boot, conf_level = 0.95) 
y_pred_cpp <- results$y_pred
lower_cpp <- results$lower
upper_cpp <- results$upper

# Predict with the np package (R implementation)
np_model <- npreg(bws = bandwidth, xdat = x, ydat = y, exdat = x_eval, 
                  ckertype = "gaussian")
y_pred_np <- predict(np_model)

# Plot the results with confidence intervals
plot(x, y, main = "Nonparametric Regression Comparison", xlab = "x", ylab = "y")
lines(x_eval, y_pred_cpp, col = "blue", lwd = 2, lty = 1)
lines(x_eval, lower_cpp, col = "blue", lty = 2)
lines(x_eval, upper_cpp, col = "blue", lty = 2)
lines(x_eval, y_pred_np, col = "red", lwd = 2, lty = 1)
legend("topright", legend = c("R Estimate", "R Confidence Interval", "np Package Estimate"), 
       col = c("blue", "blue", "red"), lty = c(1, 2, 1), lwd = c(2, 1, 2))
```
